Variability of Planetary Mass Companion 2M1207 B

### 2015-09-08 summary/letter/letter_vAC.tex
do remember to bring the adapter!!

\title{Variability of Planetary Mass Companion 2M1207 B}
\shorttitle{Variability of 2M1207 B}
\author{Yifan Zhou, Daniel Apai, Glenn Schneider, ...}
\affil{The University of Arizona}

@@

The observations applied space craft rolls between each two orbits to
allow roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbits 1, 3, and 5, and those taken in
orbits 2, 4, and 6 differed by $25^{\circ}$. At the separation of
2M1207b this angle difference corresponds to a displacement of
$0.34''$ or 2.75 and 2.30 resolution elements in F125W and F160W,
respectively. In each orbit we took thirteen SPARS10 non-destructive
read-outs with NSAMP=10, alternating between F160W and F125W filters,
with 2--3 identical exposures in one exposure sequence. To improve PSF
sampling and reduce the risk caused by bad pixels, we applied standard
4 point dithering. Over the 6 orbits, we obtained 70 images with 10
non-destructive read-outs in F125W and 64 images in F160W with
exposure time of 88.4~s in each filters.

@@

\begin{figure*}
  \centering
  \plottwo{original}{subtracted}
  \caption{WFC3 F160W images of 2M1207A system. {\em Upper panel}: In the original image 
    2M1207 B is hidden in the halo of the bright 2M1207A.   {\em Lower panel:} In the residual image -- after the subtraction of the hybrid PSF -- 2M1207 B is detected at a high significant level.}
  \label{fig:1}
\end{figure*}

@@

We started the reduction from the the \flt files produced by the
WFC3's \texttt{calwfc3} pipeline. We did not opt to use the more
advance \ima pipeline products because these provided less information
on 2M1207A, which saturated after the first few read-outs.  The \flt
are results of basic calibration, including dark current correction,
non-linearity correction, flat field correction, as well as
up-the-ramp fit on the non-destructive read-outs also considering
cosmic rays. From the beginning, pixels with data quality flags ``bad
detector pixels'' (DQ = 4), ``unstable response'' (DQ = 32), and ``bad
or uncertain flat value'' (DQ = 512) were masked out and excluded from
further analysis as suggested by previous transit exoplanet
spectroscopic observations\citep[e.g.][]{Berta2012, Kreidberg2014}.

One major challenge of high contrast imaging observation using WFC3/IR
is significant under-sampling of the detector.  2M1207 A and b are
only separated by $\sim6$ pixels or $\sim$5 FWHM of the PSF. When
applying roll subtraction, notable artifact structures are generated
by image interpolation and shifting. On the other hand, \tinytim{} PSF
simulator\citep{Krist1995} offers a solution by providing Nyquist
sampled PSF, but significant systematic errors of \tinytim{} PSF
 limits its ability in high precision
photometry\citep{Biretta2014}. However, 6 orbits of observation
provides us an opportunity to
fully characterize the difference of model and observed PSFs. To obtain
robust \tinytim{} PSF photometry,  we designed a 2-round PSF fitting strategy: 1. calculating
correction map for \tinytim{}; 2. hybrid PSF photometry.

For both of 2 rounds, we used {\em Tiny Tim} to calculate 10$\times$
over-sampled model PSFs based the filters, the spectra 
\citep{Bonnefoy2014, Patience2010}, the telescope's actual focus, and
the telescope jitter.  We used the new set of Tiny Tim parameters
provided by \cite{Biretta2014} to better model the cold mask,
diffraction spikes, and the coma. The focus parameters were calculated
using the model listed on the STScI
website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}.
We determined the exact position of 2M1207A first by fitting the PSF
on a dynamic coordinate grid and minimizing the difference in the
observed and modeled PSFs at a region centered on 2M1207 A with a
5-pixel-radius aperture centered on 2M1207 b excluded. Then introduce
another \tinytim{} PSF for 2M1207b and  fit the
position of 2M1207 b and the photometries of 2M1207 A and b together
by least square optimization. From our data set, we discovered that
the difference of observed PSFs and model PSFs are very stable for
given PSF position. Therefore at the end of the first round PSF
fitting, we derived 8 (2 roll angles $\times$ 4 dithering positions)
empirical correction maps for each filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{model} - PSF_{obs} )}
\end{equation}
where $\mathrm{PSF_{model}}$ is a combination of two \tinytim{} PSFs
for 2M1207 A and b. In the second round, we combined the correction
term linearly with the two \tinytim{} PSFs to generate hybrid PSFs,
and scaled the three component together to get best fit. We found that by including the correction term,
the reduced $\chi^{2}$ of PSF fitting is decreased from $\sim 10$ to
around unity. Because the total fluxes of the model PSFs were
normalized to unity as default, the fluxes of 2M1207 A and B were
solely represented by the amplitude of the two PSFs.

However, PSF profiles change with different exposure positions due to
pixelation, especially for the case that WFC3 IR is significantly
under-sampled. Also, the flat fields may potentially have large scale
structures \citep{dressel2012wide}. Because of these factors, the
fluxes for both 2M1207 A and B are correlated with the dithering
position. To break the correlation, we normalize each group of
exposures that have the same dithering position and roll angle
individually -- we took the median of the fluxes that were measured
from these exposures as normalization factors and devided them from
every photometric measurement. Because the normalization factor for
each group of exposures is calculated across the whole observation,
this normalization step have negligible impacts on variability
analysis.

@@

% %\section{Uncertainty analysis}

\subsection{Uncertainty Analysis: White noise}
We consider the primary source of uncertainty for every measurement is
photon noise. We propagated the photon noises of every single pixel
that were calculated from count rates and detector gain, to the PSF
fitting results. The photon noises for photometry in F125W and F160W
are 1.33\% and 1.02\%, respectively.

Since the primary and secondary PSFs were fitted simultaneously in our
PSF photometry procedure, the uncertainties of photometry and position
for the primary and secondary are coupled with each other. The second
order uncertainty could also contribute to the overall uncertainty,
e.g. the imperfection of position measurement of 2M1207A can affect
the photometry of 2M1207b. We used a Monte Carlo (MC) method to evaluate the complete systematic of the
PSF fitting. We applied the PSF photometry to images that
were added with random Poisson noise and repeat the procedure for 1000
time. From the distribution of the result, the uncertainty for F125W
and F160W photometry are found to be 1.34\% and 1.12\%,
respectively. We conclude that the white noise of our observation is
dominated by photon noise.

@@

 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
  variability verification tests. Individual measurements are plotted with gray
  crosses. Photometries of the same exposure sequence are binned, and
  binned data are plotted with points or squares. Fitted sinusoidal
  waves are plotted with solid lines. {\em Upper}: binned photometries taken
in dithering position 1 and 3 (red points) and that taken in 2 and 4
(blue squares) are plotted differently. They demonstrate same trend of
variation. In upper left panel, sinusoidal wave fitted with all
parameter set free, and that fitted with period set the same as that
of F125W are plotted with green and purple lines, respectively. {\em
  Middle}: sinusoidal waves fitted without using the data taken in
Orbit 1 are plotted. These curves are almost identical to the curves
plotted in upper panel. {\em Lower}: photometry measured with
AFEM-applied images and their best fitted sinusoidal curves are
plotted. These photometries and curves are also almost identical to
those plotted in upper panel.}
  \label{fig:2}
\end{figure*}

In our observations 2M1207 B were observed at 8 different spots on the
detector (2 rolls $\times$ 4 dithering positions). Imperfect flat
field correction are potential sources of variation as they may
introduce position-dependent differences in the count rates. The
uncertainty of WFC3 IR pipeline flat field is $\sim 1\%$
\citep{dressel2012wide}.
% and our time resolved observations for another target
%demonstrate that aperture photometry for exposures taken at different
%dithering position are uncertain at the same level.
%Flat field uncertainties have smaller effects on 
In PSF photometry, however, multiple pixels are fitted simultaneously
and we expect a lower than 1\% uncertainty from the flat field
errors. To verify this we multiplied every image by an artificial flat
field error mask (AFEM) -- a uniformly distributed Gaussian noise array with
mean of 1 and sigma of 1\% -- and repeated the PSF photometry on the
resulting images.  The analysis of these experiments showed almost
identical lightcurve to the original, verifying that the flat field
errors do not affect our photometry significantly (Figure
\ref{fig:2}).

>>>>

\title{Variability of Planetary Mass Companion 2M1207 B}
\shorttitle{Variability of 2M1207 B}
\author{Yifan Zhou, Daniel Apai, Glenn Schneider}
\affil{Steward Observatory, the University of Arizona}
\author{Mark Marley}
\affil{NASA Ames Research Center}

@@

The observations applied space craft rolls between each two orbits to
allow roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbits 1, 3, and 5, and those taken in
orbits 2, 4, and 6 differed by $25^{\circ}$. At the separation of
2M1207b this angle difference corresponds to a displacement of
$0.34''$ or 2.75 and 2.30 resolution elements in F125W and F160W,
respectively. In each orbit we took thirteen SPARS10 non-destructive
read-outs with NSAMP=10, alternating between F160W and F125W filters,
with 2--3 identical exposures in one exposure sequence. To improve
sampling and reduce the risk that point spread funciton (PSF) lays on
bad pixels, we applied a 4-point dither patternwith differential "X/Y"
offsets of 1.375" in the detector frame, providing optimal
non-integral (half pixel) stepping of 10.5 and 8.5 pixels in F125W and
F160W, respectively.. Over the 6 orbits, we obtained 70 images with 10
non-destructive read-outs in F125W and 64 images in F160W with
exposure time of 88.4~s in each filters.

@@

\begin{figure*}
  \centering
  \plottwo{original}{subtracted}
  \caption{WFC3 F160W images for 2M1207 system. {\em Upper panel}:
    original image, the position of 2M1207b is indicated using white
    circle.   {\em Lower panel:} residual image -- after the
    subtraction of the hybrid PSF, 2M1207 B is detected at a high significant level.}
  \label{fig:1}
\end{figure*}

@@

We started the reduction from the the \flt files produced by the
WFC3's \texttt{calwfc3} pipeline. We did not opt to use the more
advance \ima pipeline products because these provided less information
on 2M1207A, which saturated after the first few read-outs.  The \flt
are results of basic calibration, including dark current correction,
non-linearity correction, flat field correction, as well as
up-the-ramp fit on the non-destructive read-outs also considering
cosmic rays. From the beginning, pixels with data quality flags ``bad
detector pixels'', ``unstable response'', and ``bad
or uncertain flat value'' were masked out and excluded from
further analysis as suggested by previous transit exoplanet
spectroscopic observations\citep[e.g.][]{Berta2012, Kreidberg2014}.

One major challenge of high contrast imaging observation using WFC3/IR
is significant under-sampling of the detector.  2M1207 A and b are
only separated by $\sim6$ pixels or $\sim$5 FWHM of the PSF. When
applying roll subtraction, notable artifact structures are generated
by image interpolation and shifting. On the other hand, \tinytim{} PSF
simulator\citep{Krist1995} offers a solution by providing Nyquist or better
sampled PSF, but systematic errors of \tinytim{} PSF for WFC3 limits
its ability in high precision photometry\citep{Biretta2014}. However,
we are able to fully characterize the difference of model and observed
PSFs with 6 orbits time-resolved observation data. To obtain robust \tinytim{} PSF
photometry, we designed a 2-round PSF fitting strategy: 1. calculating
correction map for \tinytim{}; 2. hybrid PSF photometry.

For both of 2 rounds, we used {\em Tiny Tim} to calculate 10$\times$
over-sampled model PSFs based on the filters, the spectra 
\citep{Bonnefoy2014, Patience2010}, the telescope's actual focus, and
the telescope jitter.  We used the new set of Tiny Tim parameters
provided by \cite{Biretta2014} to improve model of the cold mask,
OTA spikes, and the coma. The focus parameters were calculated
using the model listed on the STScI
website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}. To
register \tinytim{} PSF to observed PSF of 2M1207A, we moved the
over-sampled PSF over a coordinate grid (gird size=0.001 pixel) using
cubic interpolation and minimize the rms difference of the
observed and re-binned \tinytim{} PSFs over a region centered on 2M1207A with a
5-pixel-radius aperture centered on 2M1207b excluded. Then we introduce
another \tinytim{} PSF for 2M1207b and  fit the
position of 2M1207 b and the photometries of 2M1207 A and b simultaneously
by least square optimization. From our data set, we discovered that
the difference of observed PSFs and model PSFs were very stable for
given PSF positions. Therefore at the end of the first round PSF
fitting, we derived 8 (2 roll angles $\times$ 4 dithering positions)
observationally derived correction maps for each filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{obs.} - PSF_{model} )}
\end{equation}
where $\mathrm{PSF_{model}}$ is a combination of two \tinytim{} PSFs
for 2M1207 A and b. In the second round, we combined the correction
term linearly with the two \tinytim{} PSFs to generate hybrid PSFs,
and fitted the three components together. We found that by including the correction term,
the reduced $\chi^{2}$ of PSF fitting is decreased from $\sim 10$ to
around unity. Relative photometry is acquired from the scaling
parameters of the model PSFs.

PSF profiles change with exposure positions due to
pixelation, especially for the case that WFC3 IR is significantly
under-sampled. Also, the flat fields may potentially have large scale
structures \citep{dressel2012wide}. We found a correlation of
photometry with PSF positions on detector frame for both 2M1207 A and
b. Correction were made by normalizing each group of
exposures that have the same dithering position and roll angle
individually -- we took the median of the fluxes that were measured
from these exposures as normalization factors and divided them from
every photometric measurement. Because the normalization factor for
each group of exposures is calculated across the whole observation,
this normalization step have negligible impacts on variability
analysis.

@@

 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
  variability verification tests. Individual measurements are plotted with gray
  crosses. Photometries of the same exposure sequence are binned, and
  binned photometry are  plotted with points or squares. Best fitted sinusoidal
  curves are plotted with solid lines. {\em Upper}: binned photometries taken
in dithering position 1 and 3 (red points) and that taken in 2 and 4
(blue squares) are plotted with different symbols. They demonstrate same trend of
variation. In upper left panel, green line is sinusoidal wave fitted with all
parameter set free, and purple line is sinusoidal wave fitted with period set the same as that
of F125W. {\em
  Middle}: sinusoidal waves fitted without using the data taken in
Orbit \#1. These curves are almost identical to the curves
plotted in upper panel. {\em Lower}: photometry measured with
AFEM-added images and best fitted sinusoidal curves. These photometries and curves are also almost identical to
those plotted in the upper panel.}
  \label{fig:2}
\end{figure*}

\subsection{Uncertainty Analysis: White noise}
First we estimated the photon noises for the photometry of 2M1207b. We
propagated the photon noises of every single pixel that were
calculated from count rates and detector gain, to the PSF fitting
results. The photon noises for photometry in F125W and F160W are
1.33\% and 1.02\%, respectively.

Since the PSFs for the 2M1207 A and b were fitted simultaneously in our
PSF photometry procedure, the uncertainties of photometry and position
for the primary and secondary are coupled. Imperfection of position
measurement of 2M1207A could potentially affect the photometry of
2M1207b. We used a Monte Carlo (MC) method to
evaluate the overall systematic of the PSF fitting. We applied the
PSF photometry to images that were added with random Poisson noises and
repeat the photometry procedure for 1000 times. From the distribution of the
result, the uncertainty for F125W and F160W photometry are 
1.34\% and 1.12\%, respectively. We conclude that the white noise of
our observation is dominated by photon noise.

@@

2M1207b were observed at 8 different spots on the
detector (2 rolls $\times$ 4 dithering positions). Imperfect flat
field correction are potential sources of variation as they may
introduce position-dependent differences in the count rates. The
uncertainty of WFC3 IR pipeline flat field is typically $\sim 1\%$
\citep{dressel2012wide}.
% and our time resolved observations for another target
%demonstrate that aperture photometry for exposures taken at different
%dithering position are uncertain at the same level.
%Flat field uncertainties have smaller effects on 
In PSF photometry, however, multiple pixels are fitted simultaneously
and we expect a lower than 1\% uncertainty from the flat field
errors. To verify this we multiplied every image by an artificial flat
field error mask (AFEM) -- a uniformly distributed Gaussian noise array with
mean of 1 and sigma of 1\% -- and repeated the PSF photometry on the
resulting images.  The analysis of these experiments showed almost
identical lightcurve to the original, verifying that the flat field
errors do not affect our photometry significantly (Figure
\ref{fig:2}).
